Supplementary material S3: Biodiversity net effect and resilience

### 2018-11-06 S2-Resilience.Rmd
# Draft 0.1.9007 * Fourth version after @BrunoHerault meeting

---
title: "Supplementary material S2: Resilience indice"
date: '`r Sys.Date()`'
output:
  bookdown::html_document2:
    number_sections: false
    toc: true
    toc_float: yes
  bookdown::pdf_document2:
    number_sections: false
    toc: false
csl: /home/sylvain/Documents/Bibliography/csl/mee.csl
bibliography: /home/sylvain/Documents/Bibliography/library.bib
link-citations: yes
---

@@

```{r IrGraph, eval=F, include=F}
mature %>% 
  select(time, sim, agb) %>% 
  separate(sim, c("SR", "CVH"), remove = F) %>% 
  mutate(disturbance = 50) %>% 
  rbind(ecosystem %>% 
          select(time, sim, agb) %>% 
          separate(sim, c("SR", "CVH", "disturbance"), remove = F)) %>% 
  filter(SR == 25, CVH == 4, disturbance %in% c(0, 50)) %>% 
  reshape2::dcast(time ~ disturbance, value.var = "agb") %>% 
  rename(control = `0`, treatment = `50`) %>% 
  ggplot(aes(x = time)) +
  geom_line(aes(y = control, color = "control")) +
  geom_line(aes(y = treatment, color = "disturbed")) +
  geom_ribbon(aes(ymax = control, ymin = treatment), 
              fill = "lightgrey", col = NA, alpha = 0.5) +
  xlab("Time (years)") +
  ylab("Aboveground biomass (tonC/ha)") +
  scale_color_discrete("Simulation")
```

This document introduce and explain the recovery function $R(X)$ and the resilience indice $I_R$ definitions (figure \@ref(fig:IrGraphFinal)).

@@

The recovery of the simulated forest ecosystem at a time $t$, $R(t)$ regarding a given properties $X$ was defined ad follows:

@@

When $R(t)$ approaches $R_{eq}=1$, the disturbed and control systems are considered equivalent, indicating a perfect recovery. We then calculated the euclidean distance to this state of perfect recovery as $d_{eq}(t) = \sqrt{(R_{eq}-R(t))^2}$. Because the intensity of disturbance was the same for all simulations, our analysis did not include any ecosystem resistance. Consequently, we can use the cummulated recovery as an indice of ecosystem resilience, further defined as $I_R$:

@@

Ecosystem euclidean distance to equilibrium distance was calculated in a multi-dimensional space for the two following forest functions : (i) forest structure (AGB, BA, N, N10, and N30) and (ii) forest functioning (GPP, NPP, Rday, and Rnight). We then integrated the euclidean distances over time to get our ecosystem resilience indice. (Figure \@ref(fig:datatrans)). We used the cumulative integral after 600 years defined as $I_R$ as a measurement of ecosystem resilience. A high value of $I_R$ indicating a slow recovery.

@@

```{r}

```

>>>>

---
title: "Supplementary material S2: Resilience indice"
date: '`r Sys.Date()`'
output:
  bookdown::html_document2:
    number_sections: false
    toc: true
    toc_float: yes
  bookdown::pdf_document2:
    number_sections: false
    toc: false
  bookdown::word_document2: default
csl: /home/sylvain/Documents/Bibliography/csl/mee.csl
bibliography: /home/sylvain/Documents/Bibliography/library.bib
link-citations: yes
---

@@

```{r IrGraph, eval=F, include=F}
mature %>% 
  select(time, sim, agb) %>% 
  separate(sim, c("SR", "CVH"), remove = F) %>% 
  mutate(disturbance = 50) %>% 
  rbind(ecosystem %>% 
          select(time, sim, agb) %>% 
          separate(sim, c("SR", "CVH", "disturbance"), remove = F)) %>% 
  filter(SR == 25, CVH == 4, disturbance %in% c(0, 50)) %>% 
  reshape2::dcast(time ~ disturbance, value.var = "agb") %>% 
  rename(control = `0`, treatment = `50`) %>% 
  ggplot(aes(x = time)) +
  geom_line(aes(y = control, color = "control")) +
  geom_line(aes(y = treatment, color = "disturbed")) +
  geom_ribbon(aes(ymax = control, ymin = treatment), 
              fill = "lightgrey", col = NA, alpha = 0.5) +
  xlab("Time (years)") +
  ylab("Aboveground biomass (tonC/ha)") +
  scale_color_discrete("Simulation") +
  scale_color_discrete("Disturbance\nintensity\n(% of BA)", 
                       labels = c("0 (control)", "50 (disturbed)")) + 
  ggsave(filename = "./images/original.png", width = 20, height = 10, units = "cm")
```

This document introduces and explains the recovery function $R(X)$ and the resilience indice $I_R$ definitions (figure \@ref(fig:IrGraphFinal)). Resilience concept has met numerous definitions, but resilience in ecology can encompass both ecosystem ability to adsorb and recover from disturbance, resistance and recovery respectively [@Gunderson2010].

@@

The recovery of the simulated forest ecosystem, i.e. the ecosystem capacity to recover from a disturbance event, at a time $t$, $R(t)$ regarding a given property $X$ was defined ad follows:

@@

When $R(t)$ approaches $R_{eq}=1$, the disturbed and control systems are considered equivalent, indicating a perfect recovery. We then calculated the euclidean distance to this state of perfect recovery as $d_{eq}(t) = \sqrt{(R_{eq}-R(t))^2}$. Because the intensity of disturbance was the same for all simulations, our analysis did not include any ecosystem resistance, i.e. the ecosystem capacity to absorb a disturbance event,. Consequently **in this particular case**, we can use the cummulated recovery as an indice of ecosystem resilience, further defined as $I_R$:

@@

A high value of $I_R$ indicating a slow recovery. Ecosystem euclidean distance to equilibrium distance was calculated in a multi-dimensional space for the two following forest functions : (i) forest structure (AGB, BA, N, N10, and N30, e.g. figure \@ref(fig:datatrans)) and (ii) forest functioning (GPP, NPP, Rday, and Rnight).

@@

# References

### 2018-11-06 S3-RNE.Rmd
# Draft 0.1.9007 * Fourth version after @BrunoHerault meeting

We want to merge two functions to analyse the output of ecosystem after disturbance: (i) the biodiversity net effect (NE) and its partitionning in complementarity (CE) and selection (SE) effects, and (ii) the resilience function of the ecosystem. More particularly we computed $R[NE(X)]$, and we want to know if $R[NE(X)]=NE[R(X)]$. In order to do that we will proceed in three times. First we need to define all mathematic functions. Then we will use a simulation approach to test for the equality (easy way to reject it). Finally if the simulation did not rejected the equality we will try to prove it mathematically.

@@

We first need to define the resilience function, the biodiversity net effect functions, and their composed functions for any variable $X$ from the ecosystem.

@@

Resilience has been defined by us as:

\begin{equation}
  R(t) = \frac{X_T(t)}{X_C(t)}
  (\#eq:R)
\end{equation}

where $X_T(t)$ and $X_C(t)$ are the ecosystem metric values at time $t$ in the disturbed simulation and in the undisturbed control simulation.

@@

The net effect of biodiversity on ecosystem recovery (NE) is defined as the difference between the ecosystem recovery of the mixed-species community and the one expected under the null hypothesis that there is no effect of biodiversity, i.e. ecosystem recovery for the mixed-species community equals the abundance-weighted sum of that in monoculture [@Loreau2001]. We then split this net biodiversity effect in complementarity and selection effect using @Loreau2001 partitioning:

@@

\begin{equation}
  \begin{array}{c}
    R[NE(X)]=\frac{NE_T(X)}{NE_C(X)} \\
    R[NE(X)]=\frac{CE_T(X)}{NE_C(X)} + \frac{SE_T(X)}{NE_C(X)} \\
    CE(X) = N* \overline{\Delta _R(X)} \overline{M}\\
    SE(X) = N*cov(\Delta _R(X),M) \\
    \Delta _RX_{sp} = \frac{X_{sp}(mixture)}{X_{sp}(monoculture)} - P_{sp}
  \end{array}
  (\#eq:RNE)
\end{equation}

@@

\begin{equation}
  \begin{array}{c}
    \Delta _R[R(X)]_{sp} = \frac{R(X)_{sp}(mixture)}{R(X)_{sp}(monoculture)} - P_{sp} \\
    \Delta _R[R(X)]_{sp} = \frac{ \frac{{X_{sp}}_T(mixture)}{{X_{sp}}_C(mixture)} }{       \frac{{X_{sp}}_T(monoculture)}{{X_{sp}}_C(monoculture)} } - P_{sp} \\
    R(M_{sp})= \frac{{M_{sp}}_T}{{M_{sp}}_C} \\
    CE[R(X)] = N* \overline{\Delta _R[R(X)]} \overline{R(M)}\\
    SE[R(X)] = N*cov(\Delta _R[R(X)],R(M)) \\
    NE[R(X)] = CE[R(X)] + SE[R(X)]  
  \end{array}
  (\#eq:NER)
\end{equation}

@@

Let's imagine a mixture ($C_T$) of three species $sp1$, $sp2$ and $sp3$ with an ecosystem variable $X$ evolving after a disturbance event, the corresponding three monoculture (${M_{sp1}}_T$, ${M_{sp2}}_T$, ${M_{sp3}}_T$) which encountered the same disturbance, and the corresponding control mixture ($C_C$) and three control monoculture (${M_{sp1}}_C$, ${M_{sp2}}_C$, ${M_{sp3}}_C$) which did not encoutered any disturbance.

@@

\begin{equation}
  \begin{array}{c}
    \Delta _R[R(X)]_{sp} = \frac{ \frac{{X_{sp}}_T(mixture)}{{X_{sp}}_C(mixture)} }{ \frac{{X_{sp}}_T(monoculture)}{{X_{sp}}_C(monoculture)} } - P_{sp} \\
    \Delta _R[R(X)]_{sp} = 0 \\
    \frac{ \frac{{X_{sp}}_T(mixture)}{{X_{sp}}_C(mixture)} }{ \frac{{X_{sp}}_T(monoculture)}{{X_{sp}}_C(monoculture)} } = P_{sp} \\
    \frac{{X_{sp}}_T(mixture)}{{X_{sp}}_C(mixture)} = P_{sp} * \frac{{X_{sp}}_T(monoculture)}{{X_{sp}}_C(monoculture)}\\
    \frac{{X_{sp}}_T(mixture)}{{X_{sp}}_T(monoculture)}*\frac{1}{{X_{sp}}_C(mixture)} = P_{sp} * \frac{1}{{X_{sp}}_C(monoculture)}\\
    \frac{{X_{sp}}_T(mixture)}{{X_{sp}}_T(monoculture)} = P_{sp} * \frac{{X_{sp}}_C(mixture)}{{X_{sp}}_C(monoculture)}\\
    \frac{{X_{sp}}_T(mixture)}{{X_{sp}}_T(monoculture)} + P_{sp} - P_{sp} = P_{sp} * (\frac{{X_{sp}}_C(mixture)}{{X_{sp}}_C(monoculture)} + P_{sp} - P_{sp})\\
    {\Delta _R(X)_{sp}}_T + P_{sp} = P_{sp} * ({\Delta _R(X)_{sp}}_C + P_{sp}) \\
    {\Delta _R(X)_{sp}}_T + 1 = {\Delta _R(X)_{sp}}_C + P_{sp}, \forall P_{sp} \neq 0\\
    {\Delta _R(X)_{sp}}_T = {\Delta _R(X)_{sp}}_C + P_{sp} - 1 \\
    R[\Delta _R(X)_{sp}] = 1 + \frac{ P_{sp} - 1 }{ {\Delta _R(X)_{sp}}_C }, \forall {\Delta _R(X)_{sp}}_C \neq 0
  \end{array}
  (\#eq:DeltaR0)
\end{equation}

@@

\begin{equation}
  \begin{array}{c}
    \Delta _R[R(X)]_{sp} = \frac{ \frac{{X_{sp}}_T(mixture)}{{X_{sp}}_C(mixture)} }{ \frac{{X_{sp}}_T(monoculture)}{{X_{sp}}_C(monoculture)} } - P_{sp} \\
    \Delta _R[R(X)]_{sp} = a \\
    \frac{ \frac{{X_{sp}}_T(mixture)}{{X_{sp}}_C(mixture)} }{ \frac{{X_{sp}}_T(monoculture)}{{X_{sp}}_C(monoculture)} } = (P_{sp} + a) \\
    \frac{{X_{sp}}_T(mixture)}{{X_{sp}}_C(mixture)} = (P_{sp} + a) * \frac{{X_{sp}}_T(monoculture)}{{X_{sp}}_C(monoculture)}\\
    \frac{{X_{sp}}_T(mixture)}{{X_{sp}}_T(monoculture)}*\frac{1}{{X_{sp}}_C(mixture)} = (P_{sp} + a) * \frac{1}{{X_{sp}}_C(monoculture)}\\
    \frac{{X_{sp}}_T(mixture)}{{X_{sp}}_T(monoculture)} = (P_{sp} + a) * \frac{{X_{sp}}_C(mixture)}{{X_{sp}}_C(monoculture)}\\
    \frac{{X_{sp}}_T(mixture)}{{X_{sp}}_T(monoculture)} + P_{sp} - P_{sp} = (P_{sp} + a) * (\frac{{X_{sp}}_C(mixture)}{{X_{sp}}_C(monoculture)} + P_{sp} - P_{sp})\\
    {\Delta _R(X)_{sp}}_T + P_{sp} = (P_{sp} + a) * ({\Delta _R(X)_{sp}}_C + P_{sp}) \\
    {\Delta _R(X)_{sp}}_T = (P_{sp} + a)*{\Delta _R(X)_{sp}}_C + (P_{sp} + a)*P_{sp} - P_{sp} \\
    {\Delta _R(X)_{sp}}_T = (P_{sp} + a)*{\Delta _R(X)_{sp}}_C + P_{sp}*(P_{sp} + a - 1)\\
    R[\Delta _R(X)_{sp}] = P_{sp} + a + \frac{ P_{sp}*(P_{sp} + a - 1) }{ {\Delta _R(X)_{sp}}_C } \forall {\Delta _R(X)_{sp}}_C \neq 0 \\
    R[\Delta _R(X)_{sp}] = \Delta _R[R(X)]_{sp} + P_{sp} + \frac{ P_{sp}*(P_{sp} + a - 1) }{ {\Delta _R(X)_{sp}}_C } \forall {\Delta _R(X)_{sp}}_C \neq 0
  \end{array}
  (\#eq:DeltaRa)
\end{equation}

@@

This one is straightforward, it only depends on the order of the indices.
\begin{equation}
  M[R(X)] = \frac{{X_{sp}}_T(monoculture)}{{X_{sp}}_C(monoculture)} = R[M(X)]
  (\#eq:M)
\end{equation}

@@

\begin{equation}
  \begin{array}{c}
    CE[R(X)] = N* \overline{\Delta _R[R(X)]} \overline{R(M)}\\
    R[\Delta _R(X)_{sp}] = \Delta _R[R(X)]_{sp} + P_{sp} + \frac{ P_{sp}*(P_{sp} + a - 1) }{ {\Delta _R(X)_{sp}}_C } \forall {\Delta _R(X)_{sp}}_C \neq 0 \\
    M[R(X)] = \frac{{X_{sp}}_T(monoculture)}{{X_{sp}}_C(monoculture)} = R[M(X)]
  \end{array}
  (\#eq:CER1)
\end{equation}

@@

\begin{equation}
  \begin{array}{c}
    SE[R(X)] = N*cov(\Delta _R[R(X)],M[R(X)]) \\
    R[\Delta _R(X)_{sp}] = \Delta _R[R(X)]_{sp} + P_{sp} + \frac{ P_{sp}*(P_{sp} + a - 1) }{ {\Delta _R(X)_{sp}}_C } \forall {\Delta _R(X)_{sp}}_C \neq 0 \\
    M[R(X)] = \frac{{X_{sp}}_T(monoculture)}{{X_{sp}}_C(monoculture)} = R[M(X)]
  \end{array}
  (\#eq:SER1)
\end{equation}

>>>>

We want to merge two functions to analyse the output of ecosystem after disturbance: (i) the biodiversity net effect (NE) and its partitionning in complementarity (CE) and selection (SE) effects, and (ii) the resilience function of the ecosystem properties (R). More particularly we computed $R[NE(X)]$, and we want to know how $R[NE(X)]$ relate to $NE[R(X)]$. In order to do that we will proceed in three times. First we need to define all mathematic functions. Then we will use a simulation approach to test for the equality (easy way to reject it). Finally if the simulation did not rejected the equality we will try to prove it mathematically.

@@

We first need to define the resilience function, the biodiversity net effect function, and their composed functions for any variable $X$ of the ecosystem.

@@

We defined resilience as follows:

\begin{equation}
  R(t) = \frac{X_D(t)}{X_C(t)}
  (\#eq:R)
\end{equation}

where $X_D(t)$ and $X_C(t)$ are the ecosystem metric values at time $t$ in the disturbed simulation and in the undisturbed control simulation.

@@

The net effect of biodiversity on an ecosystem variable X (NE) is defined as the difference between the ecosystem variable of the mixed-species community and the one expected under the null hypothesis that there is no effect of biodiversity, i.e. ecosystem variable for the mixed-species community equals the abundance-weighted sum of that in monoculture [@Loreau2001]. We then split this net biodiversity effect in complementarity and selection effects using @Loreau2001 partitioning:

@@

\begin{equation}
  \begin{array}{c}
    R[NE(X)]=\frac{NE_D(X)}{NE_C(X)} \\
    R[NE(X)]=\frac{CE_D(X)}{NE_C(X)} + \frac{SE_D(X)}{NE_C(X)} \\
    CE(X) = N* \overline{\Delta _R(X)} \overline{M}\\
    SE(X) = N*cov(\Delta _R(X),M) \\
    \Delta _RX_{sp} = \frac{X_{sp}(mixture)}{X_{sp}(monoculture)} - P_{sp}
  \end{array}
  (\#eq:RNE)
\end{equation}

@@

\begin{equation}
  \begin{array}{c}
    \Delta _R[R(X)]_{sp} = \frac{R(X)_{sp}(mixture)}{R(X)_{sp}(monoculture)} - P_{sp} \\
    \Delta _R[R(X)]_{sp} = \frac{ \frac{{X_{sp}}_D(mixture)}{{X_{sp}}_C(mixture)} }{       \frac{{X_{sp}}_D(monoculture)}{{X_{sp}}_C(monoculture)} } - P_{sp} \\
    R(M_{sp})= \frac{{M_{sp}}_D}{{M_{sp}}_C} \\
    CE[R(X)] = N* \overline{\Delta _R[R(X)]} \overline{R(M)}\\
    SE[R(X)] = N*cov(\Delta _R[R(X)],R(M)) \\
    NE[R(X)] = CE[R(X)] + SE[R(X)]  
  \end{array}
  (\#eq:NER)
\end{equation}

@@

Let's imagine a mixture ($C_D$) of three species $sp1$, $sp2$ and $sp3$ with an ecosystem variable $X$ evolving after a disturbance event, the corresponding three monoculture (${M_{sp1}}_D$, ${M_{sp2}}_D$, ${M_{sp3}}_D$) which encountered the same disturbance, and the corresponding control mixture ($C_C$) and three control monoculture (${M_{sp1}}_C$, ${M_{sp2}}_C$, ${M_{sp3}}_C$) which did not encoutered any disturbance.

@@

\begin{equation}
  \begin{array}{c}
    \Delta _R[R(X)]_{sp} = \frac{ \frac{{X_{sp}}_D(mixture)}{{X_{sp}}_C(mixture)} }{ \frac{{X_{sp}}_D(monoculture)}{{X_{sp}}_C(monoculture)} } - P_{sp} \\
    \Delta _R[R(X)]_{sp} = 0 \\
    \frac{ \frac{{X_{sp}}_D(mixture)}{{X_{sp}}_C(mixture)} }{ \frac{{X_{sp}}_D(monoculture)}{{X_{sp}}_C(monoculture)} } = P_{sp} \\
    \frac{{X_{sp}}_D(mixture)}{{X_{sp}}_C(mixture)} = P_{sp} * \frac{{X_{sp}}_D(monoculture)}{{X_{sp}}_C(monoculture)}\\
    \frac{{X_{sp}}_D(mixture)}{{X_{sp}}_D(monoculture)}*\frac{1}{{X_{sp}}_C(mixture)} = P_{sp} * \frac{1}{{X_{sp}}_C(monoculture)}\\
    \frac{{X_{sp}}_D(mixture)}{{X_{sp}}_D(monoculture)} = P_{sp} * \frac{{X_{sp}}_C(mixture)}{{X_{sp}}_C(monoculture)}\\
    \frac{{X_{sp}}_D(mixture)}{{X_{sp}}_D(monoculture)} + P_{sp} - P_{sp} = P_{sp} * (\frac{{X_{sp}}_C(mixture)}{{X_{sp}}_C(monoculture)} + P_{sp} - P_{sp})\\
    {\Delta _R(X)_{sp}}_D + P_{sp} = P_{sp} * ({\Delta _R(X)_{sp}}_C + P_{sp}) \\
    {\Delta _R(X)_{sp}}_D + 1 = {\Delta _R(X)_{sp}}_C + P_{sp}, \forall P_{sp} \neq 0\\
    {\Delta _R(X)_{sp}}_D = {\Delta _R(X)_{sp}}_C + P_{sp} - 1 \\
    R[\Delta _R(X)_{sp}] = 1 + \frac{ P_{sp} - 1 }{ {\Delta _R(X)_{sp}}_C }, \forall {\Delta _R(X)_{sp}}_C \neq 0
  \end{array}
  (\#eq:DeltaR0)
\end{equation}

@@

\begin{equation}
  \begin{array}{c}
    \Delta _R[R(X)]_{sp} = \frac{ \frac{{X_{sp}}_D(mixture)}{{X_{sp}}_C(mixture)} }{ \frac{{X_{sp}}_D(monoculture)}{{X_{sp}}_C(monoculture)} } - P_{sp} \\
    \Delta _R[R(X)]_{sp} = a \\
    \frac{ \frac{{X_{sp}}_D(mixture)}{{X_{sp}}_C(mixture)} }{ \frac{{X_{sp}}_D(monoculture)}{{X_{sp}}_C(monoculture)} } = (P_{sp} + a) \\
    \frac{{X_{sp}}_D(mixture)}{{X_{sp}}_C(mixture)} = (P_{sp} + a) * \frac{{X_{sp}}_D(monoculture)}{{X_{sp}}_C(monoculture)}\\
    \frac{{X_{sp}}_D(mixture)}{{X_{sp}}_D(monoculture)}*\frac{1}{{X_{sp}}_C(mixture)} = (P_{sp} + a) * \frac{1}{{X_{sp}}_C(monoculture)}\\
    \frac{{X_{sp}}_D(mixture)}{{X_{sp}}_D(monoculture)} = (P_{sp} + a) * \frac{{X_{sp}}_C(mixture)}{{X_{sp}}_C(monoculture)}\\
    \frac{{X_{sp}}_D(mixture)}{{X_{sp}}_D(monoculture)} + P_{sp} - P_{sp} = (P_{sp} + a) * (\frac{{X_{sp}}_C(mixture)}{{X_{sp}}_C(monoculture)} + P_{sp} - P_{sp})\\
    {\Delta _R(X)_{sp}}_D + P_{sp} = (P_{sp} + a) * ({\Delta _R(X)_{sp}}_C + P_{sp}) \\
    {\Delta _R(X)_{sp}}_D = (P_{sp} + a)*{\Delta _R(X)_{sp}}_C + (P_{sp} + a)*P_{sp} - P_{sp} \\
    {\Delta _R(X)_{sp}}_D = (P_{sp} + a)*{\Delta _R(X)_{sp}}_C + P_{sp}*(P_{sp} + a - 1)\\
    R[\Delta _R(X)_{sp}] = P_{sp} + a + \frac{ P_{sp}*(P_{sp} + a - 1) }{ {\Delta _R(X)_{sp}}_C } \forall {\Delta _R(X)_{sp}}_C \neq 0 \\
    R[\Delta _R(X)_{sp}] = \Delta _R[R(X)]_{sp} + P_{sp} + \frac{ P_{sp}*(P_{sp} + a - 1) }{ {\Delta _R(X)_{sp}}_C } \forall {\Delta _R(X)_{sp}}_C \neq 0
  \end{array}
  (\#eq:DeltaRa)
\end{equation}

@@

This one is straightforward, it only depends on the order of the indices.
\begin{equation}
  M[R(X)] = \frac{{X_{sp}}_D(monoculture)}{{X_{sp}}_C(monoculture)} = R[M(X)]
  (\#eq:M)
\end{equation}

@@

\begin{equation}
  \begin{array}{c}
    CE[R(X)] = N* \overline{\Delta _R[R(X)]} \overline{R(M)}\\
    R[\Delta _R(X)_{sp}] = \Delta _R[R(X)]_{sp} + P_{sp} + \frac{ P_{sp}*(P_{sp} + a - 1) }{ {\Delta _R(X)_{sp}}_C } \forall {\Delta _R(X)_{sp}}_C \neq 0 \\
    M[R(X)] = \frac{{X_{sp}}_D(monoculture)}{{X_{sp}}_C(monoculture)} = R[M(X)]
  \end{array}
  (\#eq:CER1)
\end{equation}

@@

\begin{equation}
  \begin{array}{c}
    SE[R(X)] = N*cov(\Delta _R[R(X)],M[R(X)]) \\
    R[\Delta _R(X)_{sp}] = \Delta _R[R(X)]_{sp} + P_{sp} + \frac{ P_{sp}*(P_{sp} + a - 1) }{ {\Delta _R(X)_{sp}}_C } \forall {\Delta _R(X)_{sp}}_C \neq 0 \\
    M[R(X)] = \frac{{X_{sp}}_D(monoculture)}{{X_{sp}}_C(monoculture)} = R[M(X)]
  \end{array}
  (\#eq:SER1)
\end{equation}
